---
title: "RWorksheet_Malayas#4c"
author: "Andrew Miguel M. Malayas BSIT2A"
date: "2024-11-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg

a. Show your solutions on how to import a csv file into the environment.
```{r}
library(readr)
mpg <- read_csv("mpg.csv")
mpg
```
b. Which variables from mpg dataset are categorical?

manufacturer - The car manufacturer (e.g., Audi, Chevrolet).
 - model
 - trans
 - drv
 - cyl
 - fl
 - class
 - manufacturer
  

c. Which are continuous variables?
 - displ
 - year
 - cyt
 - hwy
 
 
2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer.
```{r}
library(dplyr)

manufacturer_model_count <- mpg %>%
  group_by(manufacturer) %>%
  summarise(model_count = n_distinct(model)) %>%
  arrange(desc(model_count))

most_models_manufacturer <- manufacturer_model_count[1, ]

model_variation_count <- mpg %>%
  group_by(model) %>%
  summarise(variation_count = n()) %>%
  arrange(desc(variation_count))

most_variations_model <- model_variation_count[1, ]

most_models_manufacturer
most_variations_model
```
a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
unique_models <- mpg %>%
  group_by(manufacturer) %>%
  summarise(unique_models = list(unique(model))) %>%
  arrange(manufacturer)

print(unique_models)
```
b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
model <- mpg %>%
  group_by(manufacturer) %>%
  summarise(unique_count = n_distinct(model)) %>%
  arrange(desc(unique_count))


barplot(model$unique_count, 
        names.arg = model$manufacturer, 
        las = 2, 
        col = "lightgrey", 
        main = "Number of Unique Models by Manufacturer",
        xlab = "Manufacturer", 
        ylab = "Number of Unique Models")


library(ggplot2)

ggplot(model, aes(x = reorder(manufacturer, unique_count), y = unique_count)) +
  geom_bar(stat = "identity", fill = "lightgrey") +
  coord_flip() +
  labs(title = "Number of Unique Models by Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal()

```
2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
x-axis: represents the model.
y-axis: represents the manufacturer.
This code generates a scatter plot where each point represents a unique pairing of model and manufacturer. It illustrates which models are produced by each manufacturer, showing the distribution of models across different manufacturers.



b. For you, is it useful? If not, how could you modify the data to make it more informative?
Usefulness:

As it stands, the plot isn’t particularly helpful for analyzing the relationship between model and manufacturer because of the overplotting of points and the absence of numerical or continuous variables that could offer deeper insights.



3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and its results.
```{r}
Top_20 <- mpg %>%
  arrange(desc(cty)) %>%
  head(20)
ggplot(Top_20, aes(x = model, y = year)) +
  geom_point() +
  labs(title = "Top 20 Models by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5))

```


