---
title: "RWorkSheet_Malayas#2"
author: "Andrew Miguel Malayas"
date: "2024-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using Vectors
#1. Create a vector using : operator
a. Sequence from -5 to 5. Write the R code and its output.
Describe its output.

```{r}
num <- -5:5
num
```

b. x <- 1:7. What will be the value of x?
```{r}
x <- 1:7
x
```






#2.* Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
Write the R script and its output. Describe the output.
```{r}
seq(1, 3, by=0.2)
seq
```





#3. A factory has a census of its workers. There are 50 workers in total. The following list shows their ages: 
34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

```{r}
Workers <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
```

a. Access 3rd element, what is the value?
```{r}
Workers <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
Work <- Workers [3]
Work
```
b. Access 2nd and 4th element, what are the values?
```{r}
Workers <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
Work <- Workers [c(2,4)]
Work
```
c. Access all but the 4th and 12th element is not
included. Write the R script and its output.
```{r}
Workers <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18)
Work <- Workers [c(2,4)]
Work
```





#4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the
vector, names(x).

a. Print the results. Then access x[c("first", "third")].
Describe the output.

b. Write the code and its output.
```{r}
names <- c("first"=3, "second"=0, "third"=9)
ngalan <- names[c("first", "third")]
ngalan
```



#5. Create a sequence x from -3:2.
a. Modify 2nd element and change it to 0;
x[2] <- 0
x

Describe the output.
It change the value -2 to 0

b. Write the code and its output.
```{r}
x <- c(-3:2)
x[2] <- 0
x
```


#6. *The following data shows the diesel fuel purchased by Mr. Cruz.
Month Jan Feb March Apr May June
Price per liter (PhP) 52.50 57.25 60.00 65.00 74.25 54.00
Purchase–quantity(Liters) 25 30 40 50 10 45

a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.
```{r}
Month <- c("Jan", "Feb", "March", "Apr", "May", "June")
Php <- c( 52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
Liters <- c( 25, 30, 40, 50, 10, 45)
Data <- data.frame(Month, Php, Liters)
Data
```


b. What is the average fuel expenditure of Mr. Cruz from
Jan to June? Note: Use ‘weighted.mean(liter,
purchase)‘. Write the R scripts and its output.
```{r}
Month <- c("Jan", "Feb", "March", "Apr", "May", "June")
Php <- c( 52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
Liters <- c( 25, 30, 40, 50, 10, 45)
Data <- data.frame(Month, Php, Liters)
Data
Average <- weighted.mean(Php, Liters)
Average
```





#7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.
a. Type “rivers” in your R console.
Create a vector data with 7
elements, containing the number of elements (length)
in rivers, their sum (sum), mean (mean),
median(median), variance(var), standard deviation(sd),
minimum (min) and maximum (max).
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers),
sd(rivers), min(rivers), max(rivers))
b. What are the results?

c. Write the R scripts and its outputs.



#8. The table below gives the 25 most powerful celebrities and their annual pay as ranked
by the editions of Forbes magazine and as listed on the Forbes.com website.

Figure 1: Forbes Ranking
a. Create vectors according to the above table.
Write the R scripts and its output.

b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. Write the
R scripts and its output.

c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?


d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.

e. Describe its output.






#9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
a. Import the excel file into your RStudio.
What is the R script?

b. How many dimensions does the dataset have?
What is the R script? WHat is its output?

c. Select columns country, neighbourhood,
price, stars, accomodation_type, and
ratings. Write the R script.

d. Save the data as **new.RData to your RStudio. Write the
R script.

e. Display the first six rows and last six rows of the
new.RData. What is the R script?






#10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.
a. Write the R scripts and its output.

5

b. Add 2 additional vegetables after the last vegetables
in the list. What is the R script and its output?

c. Add 4 additional vegetables after index 5. How many
datapoints does your vegetable list have? What is the
R script and its output?

d. Remove the vegetables in index 5, 10, and 15. How
many vegetables were left? Write the codes and its
output.


Prepared by:
JOYCE F. JAMILE
Instructor
Without ethical considerations, AI becomes a tool of chaos and harm.