---
title: "RWorksheet_Malayas#3b"
author: "Andrew Miguel M. Malayas BSIT2A"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a data frame using the table below.

a. Write the codes.

```{r}
data <- data.frame(
  Respondents = 1:20, 
  Sex = c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2),
  Fathers_Occupaton = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1),
  Persons_at_Home = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6), 
Siblings_at_School = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2),
Types_of_Houses =c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)
)
data
```



b. Describe the data. Get the structure or the summary of the data
```{r}

str(data)

summary(data)

```


c. Is the mean number of siblings attending is 5?

d. Extract the 1st two rows and then all the columns using the subsetting functions.
Write the codes and its output.


e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.

f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.

g. Select only all Males respondent that their father occupation was farmer. Write the
codes and its output.

h. Select only all females respondent that have greater than or equal to 5 number of
siblings attending school. Write the codes and its outputs.