---
title: "RWorksheet_Malayas#3b"
author: "Andrew Miguel M. Malayas BSIT2A"
date: "2024-10-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a data frame using the table below.

a. Write the codes.

```{r}
data <- data.frame(
  Respondents = 1:20, 
  Sex = c(2,2,1,2,2,2,2,2,2,2,1,2,2,2,2,2,2,2,1,2),
  Fathers_Occupaton = c(1,3,3,3,1,2,3,1,1,1,3,2,1,3,3,1,3,1,2,1),
  Persons_at_Home = c(5,7,3,8,5,9,6,7,8,4,7,5,4,7,8,8,3,11,7,6), 
Siblings_at_School = c(6,4,4,1,2,1,5,3,1,2,3,2,5,5,2,1,2,5,3,2),
Types_of_Houses =c(1,2,3,1,1,3,3,1,2,3,2,3,2,2,3,3,3,3,3,2)
)
data
```



b. Describe the data. Get the structure or the summary of the data
```{r}

str(data)
summary(data)
data

```


c. Is the mean number of siblings attending is 5?
```{r}
meanSiblings <- mean(data$Siblings_at_School)
meanSiblings
```


d. Extract the 1st two rows and then all the columns using the subsetting functions.
Write the codes and its output.
```{r}
subsetdata <- subset(data,Respondents <=2)
subsetdata
```



e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
subrows <- data[c(3,5),c(2,4)]
subrows
```

f. Select the variable types of houses then store the vector that results as types_houses.
Write the codes.
```{r}
types_houses <- data$Types_of_Houses
types_houses
```

g. Select only all Males respondent that their father occupation was farmer. Write the codes and its output.
```{r}
males_farmers <- data[data$Sex == 1 & data$Fathers_Occupaton,]
males_farmers
```


h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.

```{r}
female_siblings <- data[data$Sex == 2 & data$Siblings_at_School >= 5,]

female_siblings
```

2. Write a R program to create an empty data frame. Using the following codes:
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
a. Describe the results.

```{r}
df <- data.frame(
  Doubles = double(), 
  Characters = character(),
  Logicals = logical(),
  Factors = factor(),
  stringsAsFactors = FALSE
)

print("Structure of the empty dataframe:")
print(str(df))

```
a. Import the csv file into the R environment. Write the codes.
```{r}
library
```


b. Convert the Sex into factor using factor() function and change it into integer.[Legend:
Male = 1 and Female = 2]. Write the R codes and its output.

c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood
= 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.










