---
title: "RWorksheet_Malayas#4b.Rmd"
author: "Andrew Miguel M. Malayas BSIT2A"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

for() loop
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixT <- matrix(0, nrow = 5, ncol = 5)
for (i in 1:5) {
  matrixT[i, i] <- vectorA[i]
}
matrixT
```

2. Print the string "*" using for() function.
```{r}
n <- 5
for (i in 1:n) {
  for (j in 1:i) {
    cat(" *  ")
  }
  cat("\n")
}

```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
start <- as.integer(readline(prompt = "Enter the starting number: "))
start <- 1
a <- start
b <- 1
cat(a, b, sep = " ")

repeat {
  nxt <- a + b
  if (!is.na(nxt) && nxt > 500) {
    break
  }
 cat(nxt, " ")
  a <- b
  b <- nxt
  }
cat("\n")
```