---
title: "RWorksheet_Malayas#4b.Rmd"
author: "Andrew Miguel M. Malayas BSIT2A"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

for() loop
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in
Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute value

```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixT <- matrix(0, nrow = 5, ncol = 5)
for (i in 1:5) {
  matrixT[i, i] <- vectorA[i]
}
matrixT
```

2. Print the string "*" using for() function.
```{r}
n <- 5
for (i in 1:n) {
  for (j in 1:i) {
    cat(" *  ")
  }
  cat("\n")
}

```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
start <- as.integer(readline(prompt = "Enter the starting number: "))
start <- 1
a <- start
b <- 1
cat(a, b, sep = " ")

repeat {
  nxt <- a + b
  if (!is.na(nxt) && nxt > 500) {
    break
  }
 cat(nxt, " ")
  a <- b
  b <- nxt
  }
cat("\n")
```




4. Import the dataset as shown in Figure 1 you have created previously.
```{r}
shoes <- read.csv("shoes.csv")
```


a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the dataset? Show your codes and its result
```{r}
head(shoes)
```


b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
female_subset <- subset(shoes, Gender == "F")
male_subset <- subset(shoes, Gender == "M")

female_num <- nrow(female_subset)
male_num <- nrow(male_subset)

cat("Female:", female_num, "\n")
cat("Male:", male_num, "\n")

```


c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
gender_counts <- table(shoes$Gender)
barplot(gender_counts, 
        main = "Number of Males and Females in Household Data",
        xlab = "Gender",
        ylab = "Count",
        col = c("pink", "lightblue"),
        
        names.arg = c("Female", "Male"))
```






5. The monthly income of Dela Cruz family was spent on the following:
a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)
percent_labels <- paste0(names(expenses), " - ", round(expenses / sum(expenses) * 100), "%")
pie(expenses, 
    labels = percent_labels, 
    col = c("green", "blue", "gold", "red"), 
    main = "Dela Cruz Family Monthly Expenses")

```














